---
title: "Uncertainty Propagation and Method Comparison"
author: "Isabel Openshaw & Justin Moat"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Uncertainty Propagation and Method Comparison}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Introduction

This vignette explains how TreeCarbon handles uncertainty propagation and provides detailed guidance on comparing different allometric methods. Understanding uncertainty is crucial for carbon accounting, as it quantifies the confidence in your estimates.

**Key concepts covered:**

1. **Measurement uncertainty** - Error in DBH, height measurements
2. **Parameter uncertainty** - Uncertainty in allometric equation coefficients
3. **Model uncertainty** - Residual error from regression equations
4. **Method comparison** - Quantifying differences between allometric methods
5. **Sensitivity analysis** - How sensitive are estimates to method choice?

```{r load-package, message=FALSE, warning=FALSE}
library(TreeCarbon)
library(ggplot2)
```

---

# Part 1: Understanding Uncertainty Sources

## 1.1 Types of Uncertainty

Carbon estimates have three main sources of uncertainty:

### Measurement Uncertainty

Error in field measurements (DBH, height):

```{r measurement-unc}
# Example: DBH measurement with 2.5% error
dbh_measured <- 45  # cm
dbh_error_pct <- 2.5
dbh_sd <- dbh_measured * (dbh_error_pct / 100)

cat(sprintf("DBH: %.1f ± %.2f cm (%.1f%% error)\n", 
            dbh_measured, dbh_sd, dbh_error_pct))
cat(sprintf("95%% CI: [%.1f, %.1f] cm\n",
            dbh_measured - 1.96 * dbh_sd,
            dbh_measured + 1.96 * dbh_sd))
```

### Parameter Uncertainty

Uncertainty in allometric equation coefficients (a, b):

```{r parameter-unc}
# Bunce equation: biomass = exp(a + b * ln(π * dbh))
# Coefficients have standard errors from regression
a_mean <- -5.445  # Intercept
a_se <- 0.05      # Standard error of intercept
b_mean <- 2.505   # Slope
b_se <- 0.06      # Standard error of slope

cat("=== Parameter Uncertainty ===\n")
cat(sprintf("Intercept (a): %.3f ± %.3f\n", a_mean, a_se))
cat(sprintf("Slope (b): %.3f ± %.3f\n", b_mean, b_se))
```

### Model/Residual Uncertainty

Error from the regression model itself:

```{r model-unc}
# Residual error (model fit uncertainty)
# Typically expressed as relative error (CV)
residual_cv <- 0.10  # 10% residual CV

cat("=== Model Uncertainty ===\n")
cat(sprintf("Residual CV: %.1f%%\n", residual_cv * 100))
```

---

# Part 2: Error Propagation in TreeCarbon

## 2.1 WCC Error Propagation

The WCC method propagates error through each step of the calculation:

```{r wcc-error-prop}
# Calculate carbon with error propagation
result <- fc_agc_error(
  name = "Oak",
  dbh = 45,
  height = 18,
  type = "broadleaf",
  re_dbh = 0.025,  # 2.5% DBH measurement error
  re_h = 0.05,     # 5% height measurement error
  re = 0.025       # 2.5% coefficient error
)

cat("=== WCC Error Propagation ===\n")
cat(sprintf("Carbon: %.3f ± %.3f t C\n",
            result$AGC_WCC_t, result$sig_AGC))
cat(sprintf("CV: %.1f%%\n", 100 * result$sig_AGC / result$AGC_WCC_t))
cat(sprintf("95%% CI: [%.3f, %.3f] t C\n",
            result$AGC_WCC_t - 1.96 * result$sig_AGC,
            result$AGC_WCC_t + 1.96 * result$sig_AGC))
```

### How WCC Propagates Error

The WCC pipeline propagates uncertainty through:

1. **Tariff calculation** - Combines DBH and height errors
2. **Volume estimation** - Propagates tariff and DBH errors
3. **Biomass calculation** - Combines volume and wood density errors
4. **Carbon conversion** - Adds conversion factor uncertainty

Each step uses error propagation formulas appropriate for the mathematical operations (multiplication, addition, etc.).

## 2.2 Bunce Error Propagation

The Bunce method uses a simplified approach with CO lookup from Table 6:

```{r bunce-error-prop}
# Bunce with error propagation
result_bunce <- Bunce(
  name = "Oak",
  dbh = 45,
  re_dbh = 0.025  # Only DBH error needed
)

cat("=== Bunce Error Propagation ===\n")
cat(sprintf("Biomass: %.1f ± %.1f kg\n",
            result_bunce$biomass, result_bunce$sigma))
cat(sprintf("CV: %.1f%%\n", 
            100 * result_bunce$sigma / result_bunce$biomass))
```

### Bunce Uncertainty Method

The Bunce function uses:

1. **CO lookup** - Confidence limits from Bunce (1968) Table 6
2. **N-based adjustment** - Adjusts for sample size differences
3. **Species-specific values** - Uses species-specific CO when available
4. **Error propagation formula** - Combines measurement and parameter uncertainty

```{r bunce-method-detail}
# With rich output to see methodology
result_bunce_rich <- Bunce(
  name = "Oak",
  dbh = 45,
  re_dbh = 0.025,
  rich_output = TRUE
)

print(result_bunce_rich)
```

## 2.3 BIOMASS Monte Carlo Uncertainty

The BIOMASS package uses Monte Carlo simulation:

```{r biomass-mc, eval=requireNamespace("BIOMASS", quietly = TRUE)}
# BIOMASS with Monte Carlo uncertainty
result_biomass <- BIOMASS(
  dbh = 45,
  height = 18,
  genus = "Quercus",
  species = "robur",
  coords = c(-0.29, 51.48),
  uncertainty = TRUE,
  n_mc = 1000  # Number of Monte Carlo iterations
)

cat("=== BIOMASS Monte Carlo Uncertainty ===\n")
cat(sprintf("AGB: %.1f ± %.1f kg\n",
            result_biomass$AGB_Biomass_kg,
            result_biomass$AGB_sd_kg))
cat(sprintf("CV: %.1f%%\n", result_biomass$AGB_CV_pct))
cat(sprintf("95%% CI: [%.1f, %.1f] kg\n",
            result_biomass$AGB_CI_low_kg,
            result_biomass$AGB_CI_high_kg))
```

---

# Part 3: Comparing Uncertainty Methods

## 3.1 Visual Comparison

Compare how uncertainty varies with tree size:

```{r compare-unc-methods, fig.height=6}
# Create data for multiple DBH values
dbh_range <- seq(20, 80, by = 5)

# WCC uncertainty
wcc_results <- fc_agc_error(
  name = rep("Oak", length(dbh_range)),
  dbh = dbh_range,
  height = rep(18, length(dbh_range)),
  type = rep("broadleaf", length(dbh_range)),
  re_dbh = 0.025,
  re_h = 0.05
)

# Bunce uncertainty
bunce_results <- Bunce(
  name = rep("Oak", length(dbh_range)),
  dbh = dbh_range,
  re_dbh = 0.025
)

# Combine for plotting
plot_data <- data.frame(
  dbh = rep(dbh_range, 2),
  method = rep(c("WCC", "Bunce"), each = length(dbh_range)),
  carbon = c(wcc_results$AGC_WCC_t, bunce_results$biomass / 1000 * 0.5),
  uncertainty = c(wcc_results$sig_AGC, bunce_results$sigma / 1000 * 0.5)
)

plot_data$cv <- 100 * plot_data$uncertainty / plot_data$carbon

ggplot(plot_data, aes(x = dbh, y = cv, color = method)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 2) +
  labs(
    title = "Coefficient of Variation (CV) vs DBH",
    subtitle = "Comparing WCC and Bunce uncertainty methods",
    x = "DBH (cm)",
    y = "CV (%)",
    color = "Method"
  ) +
  theme_minimal() +
  scale_color_brewer(palette = "Set2")
```

## 3.2 Uncertainty Decomposition

For advanced users, you can decompose uncertainty by source:

```{r uncertainty-decomp, eval=requireNamespace("BIOMASS", quietly = TRUE)}
# Use mc_uncertainty for detailed decomposition
bunce_fn <- function(dbh, a, b) {
  exp(a + b * log(pi * dbh))
}

# Define inputs with uncertainties
inputs <- list(
  dbh = list(mean = 45, sd = 1.13, source = "measurement"),
  a = list(mean = -5.445, sd = 0.05, source = "parameter"),
  b = list(mean = 2.505, sd = 0.06, source = "parameter")
)

# Run Monte Carlo with decomposition
mc_result <- mc_uncertainty(
  bunce_fn,
  inputs,
  N = 2000,
  decompose = TRUE
)

# View decomposition
if (!is.null(mc_result$decomposition)) {
  cat("=== Uncertainty Decomposition ===\n")
  print(mc_result$decomposition$by_source)
}
```

---

# Part 4: Method Comparison Methodology

## 4.1 Why Compare Methods?

Different allometric methods can give different estimates for the same tree. Understanding these differences is crucial for:

- **Carbon accounting** - Reporting defensible estimates
- **Method selection** - Choosing the most appropriate method
- **Uncertainty quantification** - Understanding total uncertainty including method choice

## 4.2 Using allometries() for Comparison

The `allometries()` function runs all methods simultaneously:

```{r allometries-compare, eval=requireNamespace("BIOMASS", quietly = TRUE) && requireNamespace("allodb", quietly = TRUE)}
# Compare all methods for a single tree
comparison <- allometries(
  genus = "Quercus",
  species = "robur",
  dbh = 45,
  height = 18,
  coords = c(-0.29, 51.48),
  returnv = "AGC"
)

cat("=== Method Comparison (Single Tree) ===\n")
print(comparison[, c("genus", "species", "dbh", "height",
                      "WCC_C_t", "biomass_C_t", "allodb_C_t", "Bunce_C_t")])
```

## 4.3 Quantifying Method Differences

### Absolute Differences

```{r method-diff-abs}
# Extract estimates
estimates <- c(
  WCC = comparison$WCC_C_t,
  BIOMASS = comparison$biomass_C_t,
  allodb = comparison$allodb_C_t,
  Bunce = comparison$Bunce_C_t
)

# Remove NA values
estimates <- estimates[!is.na(estimates)]

cat("=== Absolute Differences ===\n")
cat(sprintf("Range: %.3f - %.3f t C\n", min(estimates), max(estimates)))
cat(sprintf("Spread: %.3f t C (%.1f%% of mean)\n",
            max(estimates) - min(estimates),
            100 * (max(estimates) - min(estimates)) / mean(estimates)))
cat(sprintf("Mean: %.3f t C\n", mean(estimates)))
cat(sprintf("SD: %.3f t C (CV = %.1f%%)\n",
            sd(estimates), 100 * sd(estimates) / mean(estimates)))
```

### Relative Differences

```{r method-diff-rel}
# Calculate relative differences from mean
mean_estimate <- mean(estimates)
relative_diff <- 100 * (estimates - mean_estimate) / mean_estimate

diff_table <- data.frame(
  Method = names(estimates),
  Estimate_tC = estimates,
  Difference_pct = relative_diff
)

cat("=== Relative Differences (from mean) ===\n")
print(diff_table)
```

## 4.4 Using compare_allometries()

The `compare_allometries()` function provides detailed statistics:

```{r compare-allometries-func, eval=requireNamespace("BIOMASS", quietly = TRUE) && requireNamespace("allodb", quietly = TRUE)}
# Get comparison statistics
comp_stats <- compare_allometries(
  comparison,
  reference = "WCC"  # Use WCC as reference
)

print(comp_stats)
```

---

# Part 5: Batch Method Comparison

## 5.1 Multiple Trees

Compare methods across multiple trees:

```{r batch-compare, eval=requireNamespace("BIOMASS", quietly = TRUE) && requireNamespace("allodb", quietly = TRUE)}
# Sample tree data
trees <- data.frame(
  genus = c("Quercus", "Fagus", "Fraxinus", "Betula"),
  species = c("robur", "sylvatica", "excelsior", "pendula"),
  dbh = c(45, 38, 52, 28),
  height = c(18, 22, 24, 15)
)

# Compare all methods
batch_comparison <- allometries(
  genus = trees$genus,
  species = trees$species,
  dbh = trees$dbh,
  height = trees$height,
  coords = c(-0.29, 51.48),
  returnv = "AGC"
)

cat("=== Batch Method Comparison ===\n")
print(batch_comparison[, c("genus", "species", "dbh",
                           "WCC_C_t", "biomass_C_t", "allodb_C_t", "Bunce_C_t")])
```

## 5.2 Method Agreement Statistics

```{r method-agreement}
# Calculate cross-method statistics for each tree
methods <- c("WCC_C_t", "biomass_C_t", "allodb_C_t", "Bunce_C_t")

batch_comparison$mean_estimate <- apply(
  batch_comparison[, methods], 1, mean, na.rm = TRUE
)
batch_comparison$sd_estimate <- apply(
  batch_comparison[, methods], 1, sd, na.rm = TRUE
)
batch_comparison$cv_pct <- 100 * batch_comparison$sd_estimate / batch_comparison$mean_estimate

cat("=== Method Agreement by Tree ===\n")
print(batch_comparison[, c("genus", "species", "dbh", "mean_estimate", "cv_pct")])

cat("\n=== Overall Agreement ===\n")
cat(sprintf("Mean CV across methods: %.1f%%\n",
            mean(batch_comparison$cv_pct, na.rm = TRUE)))
cat(sprintf("Range of CV: %.1f%% - %.1f%%\n",
            min(batch_comparison$cv_pct, na.rm = TRUE),
            max(batch_comparison$cv_pct, na.rm = TRUE)))
```

## 5.3 Visualizing Method Differences

```{r plot-method-diff, fig.height=6, eval=requireNamespace("BIOMASS", quietly = TRUE) && requireNamespace("allodb", quietly = TRUE)}
# Reshape data for plotting
library(tidyr)

plot_data <- batch_comparison %>%
  select(genus, species, dbh, WCC_C_t, biomass_C_t, allodb_C_t, Bunce_C_t) %>%
  pivot_longer(
    cols = c(WCC_C_t, biomass_C_t, allodb_C_t, Bunce_C_t),
    names_to = "method",
    values_to = "carbon"
  )

plot_data$method <- gsub("_C_t", "", plot_data$method)
plot_data$method <- gsub("biomass", "BIOMASS", plot_data$method)

ggplot(plot_data, aes(x = dbh, y = carbon, color = method, shape = method)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_smooth(method = "loess", se = FALSE, linewidth = 0.8) +
  labs(
    title = "Carbon Estimates by Method vs DBH",
    x = "DBH (cm)",
    y = "Above-Ground Carbon (tonnes)",
    color = "Method",
    shape = "Method"
  ) +
  theme_minimal() +
  scale_color_brewer(palette = "Set2")
```

---

# Part 6: Sensitivity Analysis

## 6.1 What is Sensitivity Analysis?

Sensitivity analysis answers: **"How much does my carbon estimate change depending on which allometric method I use?"**

This is crucial for carbon accounting because:
- Method choice can significantly affect estimates
- You need to justify your method selection
- You should report the range of possible estimates

## 6.2 Single Tree Sensitivity

```{r sensitivity-single, eval=requireNamespace("BIOMASS", quietly = TRUE) && requireNamespace("allodb", quietly = TRUE)}
# Single tree sensitivity
single_tree <- data.frame(
  dbh = 45,
  height = 18,
  genus = "Quercus",
  species = "robur"
)

# Get estimates from all methods
estimates_all <- allometries(
  genus = single_tree$genus,
  species = single_tree$species,
  dbh = single_tree$dbh,
  height = single_tree$height,
  coords = c(-0.29, 51.48),
  returnv = "AGC"
)

# Calculate sensitivity metrics
carbon_vals <- c(
  estimates_all$WCC_C_t,
  estimates_all$biomass_C_t,
  estimates_all$allodb_C_t,
  estimates_all$Bunce_C_t
)
carbon_vals <- carbon_vals[!is.na(carbon_vals)]

mean_carbon <- mean(carbon_vals)
sd_carbon <- sd(carbon_vals)
cv_carbon <- 100 * sd_carbon / mean_carbon
spread_pct <- 100 * (max(carbon_vals) - min(carbon_vals)) / mean_carbon
range_ratio <- max(carbon_vals) / min(carbon_vals)

cat("=== Single Tree Sensitivity Analysis ===\n")
cat(sprintf("Mean estimate: %.3f t C\n", mean_carbon))
cat(sprintf("SD: %.3f t C\n", sd_carbon))
cat(sprintf("CV: %.1f%%\n", cv_carbon))
cat(sprintf("Spread: %.1f%%\n", spread_pct))
cat(sprintf("Range ratio: %.2fx\n", range_ratio))
cat(sprintf("Range: %.3f - %.3f t C\n", min(carbon_vals), max(carbon_vals)))

# Interpretation
if (cv_carbon < 10) {
  interpretation <- "LOW sensitivity - method choice has minimal impact"
} else if (cv_carbon < 25) {
  interpretation <- "MODERATE sensitivity - typical variation"
} else if (cv_carbon < 50) {
  interpretation <- "HIGH sensitivity - method choice matters"
} else {
  interpretation <- "VERY HIGH sensitivity - investigate data quality"
}

cat(sprintf("\nInterpretation: %s\n", interpretation))
```

## 6.3 Batch Sensitivity Analysis

```{r sensitivity-batch, eval=requireNamespace("BIOMASS", quietly = TRUE) && requireNamespace("allodb", quietly = TRUE)}
# Multiple trees
forest_data <- data.frame(
  tree_id = 1:10,
  genus = rep(c("Quercus", "Fagus", "Fraxinus", "Betula", "Acer"), 2),
  species = rep(c("robur", "sylvatica", "excelsior", "pendula", "pseudoplatanus"), 2),
  dbh = c(45, 38, 52, 28, 42, 50, 35, 48, 40, 30),
  height = c(18, 22, 24, 15, 20, 21, 19, 23, 18, 14)
)

# Compare all methods
batch_results <- allometries(
  genus = forest_data$genus,
  species = forest_data$species,
  dbh = forest_data$dbh,
  height = forest_data$height,
  coords = c(-0.29, 51.48),
  returnv = "AGC"
)

# Calculate per-tree sensitivity
methods <- c("WCC_C_t", "biomass_C_t", "allodb_C_t", "Bunce_C_t")
batch_results$mean_est <- apply(batch_results[, methods], 1, mean, na.rm = TRUE)
batch_results$sd_est <- apply(batch_results[, methods], 1, sd, na.rm = TRUE)
batch_results$cv_pct <- 100 * batch_results$sd_est / batch_results$mean_est

cat("=== Batch Sensitivity Analysis ===\n")
cat("Per-tree CV across methods:\n")
print(batch_results[, c("genus", "species", "dbh", "cv_pct")])

cat(sprintf("\nOverall mean CV: %.1f%%\n",
            mean(batch_results$cv_pct, na.rm = TRUE)))
```

## 6.4 Sensitivity Interpretation Guide

| CV (%) | Level | Interpretation | Action |
|--------|-------|----------------|--------|
| < 10 | LOW | Method choice has minimal impact | Use any method; report single estimate |
| 10-25 | MODERATE | Typical variation between methods | Report range; justify method choice |
| 25-50 | HIGH | Significant impact of method choice | Report all methods; discuss differences |
| > 50 | VERY HIGH | Large disagreement | Investigate data quality; check method applicability |

---

# Part 7: Combining Measurement and Method Uncertainty

## 7.1 Total Uncertainty

Total uncertainty includes both:
- **Measurement uncertainty** (from field measurements)
- **Method uncertainty** (from choice of allometric equation)

```{r total-uncertainty, eval=requireNamespace("BIOMASS", quietly = TRUE) && requireNamespace("allodb", quietly = TRUE)}
# For a single tree, get estimates with uncertainty from each method
tree_dbh <- 45
tree_height <- 18

# WCC with measurement uncertainty
wcc_result <- fc_agc_error(
  name = "Oak",
  dbh = tree_dbh,
  height = tree_height,
  type = "broadleaf",
  re_dbh = 0.025,
  re_h = 0.05
)

# Bunce with measurement uncertainty
bunce_result <- Bunce(
  name = "Oak",
  dbh = tree_dbh,
  re_dbh = 0.025
)

# Method uncertainty (spread across methods)
all_methods <- allometries(
  genus = "Quercus",
  species = "robur",
  dbh = tree_dbh,
  height = tree_height,
  coords = c(-0.29, 51.48),
  returnv = "AGC"
)

method_estimates <- c(
  all_methods$WCC_C_t,
  all_methods$biomass_C_t,
  all_methods$allodb_C_t,
  all_methods$Bunce_C_t
)
method_estimates <- method_estimates[!is.na(method_estimates)]
method_uncertainty <- sd(method_estimates)

cat("=== Total Uncertainty Breakdown ===\n")
cat("WCC (with measurement error):\n")
cat(sprintf("  Estimate: %.3f ± %.3f t C\n",
            wcc_result$AGC_WCC_t, wcc_result$sig_AGC))

cat("\nBunce (with measurement error):\n")
cat(sprintf("  Estimate: %.3f ± %.3f t C\n",
            bunce_result$biomass / 1000 * 0.5,
            bunce_result$sigma / 1000 * 0.5))

cat("\nMethod uncertainty (spread across methods):\n")
cat(sprintf("  SD across methods: %.3f t C\n", method_uncertainty))
cat(sprintf("  Mean estimate: %.3f t C\n", mean(method_estimates)))

# Combined uncertainty (rough approximation)
# This is a simplified approach - for rigorous analysis, use Monte Carlo
mean_estimate <- mean(method_estimates)
measurement_unc <- mean(c(wcc_result$sig_AGC, bunce_result$sigma / 1000 * 0.5))
total_unc <- sqrt(measurement_unc^2 + method_uncertainty^2)

cat("\nCombined uncertainty (approximate):\n")
cat(sprintf("  Total SD: %.3f t C\n", total_unc))
cat(sprintf("  Total CV: %.1f%%\n", 100 * total_unc / mean_estimate))
```

## 7.2 Reporting Uncertainty

When reporting carbon estimates, include:

1. **Point estimate** - Mean or preferred method estimate
2. **Measurement uncertainty** - From field measurement errors
3. **Method uncertainty** - Range across different allometric methods
4. **Total uncertainty** - Combined (if appropriate)

Example reporting format:

```
Carbon estimate: 0.850 ± 0.085 t C (95% CI: [0.683, 1.017])
Method range: 0.720 - 0.920 t C (CV = 8.5%)
Total uncertainty includes measurement error (2.5% DBH, 5% height) 
and method choice uncertainty.
```

---

# Part 8: Best Practices

## 8.1 When to Include Uncertainty

**Always include uncertainty when:**
- Reporting for carbon markets or offsetting
- Publishing research results
- Making policy decisions
- Comparing different sites or time periods

**Uncertainty is less critical for:**
- Exploratory analysis
- Internal planning
- Rough estimates

## 8.2 Choosing Error Parameters

**DBH measurement error (re_dbh):**
- High precision: 0.01-0.02 (1-2%)
- Standard: 0.025 (2.5%)
- Low precision: 0.05-0.10 (5-10%)

**Height measurement error (re_h):**
- High precision: 0.03-0.05 (3-5%)
- Standard: 0.05-0.10 (5-10%)
- Low precision: 0.10-0.15 (10-15%)

## 8.3 Method Selection Guidelines

1. **UK temperate woodland** → Use WCC
2. **UK deciduous, no height** → Use Bunce
3. **Tropical forests** → Use BIOMASS
4. **Global extratropical** → Use allodb
5. **Multiple methods available** → Compare and report range

## 8.4 Reporting Recommendations

When reporting carbon estimates:

1. **State the method(s) used**
2. **Report uncertainty** (SD or 95% CI)
3. **Include method range** if multiple methods used
4. **Document assumptions** (measurement errors, etc.)
5. **Justify method choice** if only one method used

---

# References

1. **Error Propagation**: Chave, J., Condit, R., Aguilar, S., et al. (2004). Error propagation and scaling for tropical forest biomass estimates. *Philosophical Transactions of the Royal Society B*, 359(1443), 409-420.

2. **Method Comparison**: Réjou-Méchain, M., Tanguy, A., Piponiot, C., Chave, J., & Hérault, B. (2017). BIOMASS: An R package for estimating above-ground biomass and its uncertainty in tropical forests. *Methods in Ecology and Evolution*, 8(9), 1163-1167.

3. **Uncertainty Quantification**: Gonzalez-Akre, E., Piponiot, C., Lepore, M., et al. (2022). allodb: An R package for biomass estimation at globally distributed extratropical forest plots. *Methods in Ecology and Evolution*, 13(2), 330-338.

---

# Session Info

```{r session-info}
sessionInfo()
```
