---
title: "Woodland Carbon Code (WCC) Methods: A Complete Guide"
author: "Isabel Openshaw & Justin Moat"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Woodland Carbon Code (WCC) Methods: A Complete Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

# Introduction

The **Woodland Carbon Code (WCC)** is the UK standard for woodland carbon accounting. This vignette provides a comprehensive guide to all WCC functions in the TreeCarbon package.

## Method Reference: Which Function to Use

| Method | Functions | Input requirements | Section |
|--------|-----------|-------------------|---------|
| **A** | `tariff_vol_area()`, `merchtreevol()`, `treevol()`, `woodbiomass()`, `crownbiomass()`, `biomass2c()` | Felled sample: vol (m³), dbh (cm). Then: NSG, species for full pipeline | 2.1, 6 |
| **B** | `broadleaf_tariff()`, `tariffs()`, `fc_agc()`, `fc_agc_error()` | type = "broadleaf", species names, dbh (cm), height_timber (m) | 2.3, 7, 12.3 |
| **C** | `conifer_tariff()`, `tariffs()`, `fc_agc()`, `fc_agc_error()` | type = "conifer", species names, dbh (cm), height (total, m) | 2.2, 7, 12.3 |
| **D** | `stand_tariff()`, `calculate_top_height()`, `fc_stand_carbon()` | stand_height OR dbh + height vectors, area_ha, species/name, type | 2.4, 12.1 |
| **E** | `wcc_process_tally()`, `wcc_dbh_class_carbon()` | dbh_class, count; optional: height, area_ha, species/name, type | 12.2 |

**Notes:** Methods B/C use `fc_agc()` or `fc_agc_error()` for end-to-end carbon. Method A requires the step-by-step pipeline. Matthews2 is the default carbon conversion method (recommended for WCC).

**Runnable examples:** `6_WCC_complete_workflow.R` (full pipeline for all methods); `8_WCC_method_selection.R` (which method for your data?); `WCC_method_reference.R` (method-function-section table).

---

The WCC methodology follows this pipeline:

1. **Species lookup** - Match species names to WCC codes (Section 1)
2. **Tariff calculation** - Determine tariff number from species, height, and DBH (Section 2)
3. **Volume estimation** - Calculate merchantable and stem volumes (Section 3)
4. **Biomass calculation** - Convert volumes to biomass using wood density (NSG) (Section 4)
5. **Carbon conversion** - Convert biomass to carbon (Section 5)

```{r load-package, message=FALSE, warning=FALSE}
library(TreeCarbon)
library(ggplot2)
```

---

# 1: Species Lookup

The first step in any WCC calculation is matching your species names to WCC species codes.

## 1.1 Basic Lookup

```{r lookup-basic}
# Lookup by common name
lookupcode("Oak")
```

```{r lookup-botanical}
# Lookup by botanical name
lookupcode("Quercus robur")
```

```{r lookup-multiple}
# Lookup multiple species
species <- c("Oak", "Beech", "Scots Pine", "Ash")
lookupcode(species)
```

## 1.2 Specifying Tree Type

For ambiguous species, specify type (broadleaf or conifer):

```{r lookup-type}
# Pine could be several species - specify conifer
lookupcode("Pine", type = "conifer")
```

## 1.3 Getting Additional Information

You can get specific codes by specifying the `code` parameter:

```{r lookup-specific}
# Get NSG (wood density)
lookupcode("Oak", code = "NSG")

# Get Crown code
lookupcode("Oak", code = "Crown")

# Get Root code
lookupcode("Oak", code = "Root")
```

---

# 2: Tariff Number Calculation

The tariff number is a key parameter in WCC that relates tree form to volume. There are several ways to calculate it.

## 2.1 Method A: From Volume and Basal Area (Equation 1)

If you have measured volume from a felled sample tree (WCC Method A):

```{r tariff-vol-area}
# Example: Tree with measured volume
tariff <- tariff_vol_area(
  vol = 0.5,      # m³
  dbh = 24,       # cm
  sig_vol = 0.01, # uncertainty in volume
  re_dbh = 0.025  # 2.5% DBH measurement error
)

print(tariff)
```

## 2.2 Method C: Conifer Tariff (Equation 3)

For conifer species (WCC Method C), use species-specific equations:

```{r tariff-conifer}
# Scots Pine
tariff_pine <- conifer_tariff(
  spcode = "PS",  # Pinus sylvestris code
  height = 20,    # m
  dbh = 30,       # cm
  re_h = 0.05,    # 5% height error
  re_dbh = 0.025  # 2.5% DBH error
)

print(tariff_pine)
```

## 2.3 Method B: Broadleaf Tariff (Equation 4)

For broadleaf species (WCC Method B - timber height required):

```{r tariff-broadleaf}
# Oak
tariff_oak <- broadleaf_tariff(
  spcode = "OK",  # Quercus code
  height_timber = 15,  # m (timber height - height to first branch, WCC protocol requirement)
  dbh = 45,       # cm
  re_h = 0.05,
  re_dbh = 0.025
)

print(tariff_oak)
```

## 2.4 Method D: Stand Tariff (Equation 4)

For stand-level calculations (WCC Method D) when you have top height
(mean of 100 largest trees/ha by DBH). Use `calculate_top_height()` to
derive top height from plot data:

```{r tariff-stand}
# Stand-level tariff (Method D) - top height required
tariff_stand <- stand_tariff(
  spcode = "OK",
  stand_height = 18,  # Top height in m
  re_h = 0.05
)

print(tariff_stand)
```

## 2.5 Automatic Tariff Selection (Methods B/C)

The `tariffs()` function automatically selects broadleaf_tariff (Method B) or
conifer_tariff (Method C) based on species type:

```{r tariff-auto}
# Automatically uses conifer or broadleaf based on species
tariff_auto <- tariffs(
  spcode = "OK",  # Oak
  height = 18,
  dbh = 45,
  type = "broadleaf",
  re_h = 0.05,
  re_dbh = 0.025
)

print(tariff_auto)
```

---

# 3: Volume Estimation

## 3.1 Merchantable Volume (Equation 5a)

Merchantable volume is the volume of the main stem up to a specified top diameter:

```{r merch-vol}
# Calculate merchantable volume from tariff
merch_vol <- merchtreevol(
  dbh = 45,
  tariff = tariff_auto$tariff,
  re_dbh = 0.025,
  sig_tariff = tariff_auto$sigma
)

print(merch_vol)
```

## 3.2 Stem Volume (Equation 5b)

Stem volume includes the full stem including taper:

```{r stem-vol}
# Calculate stem volume from merchantable volume
stem_vol <- treevol(
  mtreevol = merch_vol$volume,
  dbh = 45,
  sig_mtreevol = merch_vol$sigma,
  re = 0.025
)

print(stem_vol)
```

---

# 4: Biomass Calculation

## 4.1 Stem Wood Biomass

Convert stem volume to biomass using Nominal Specific Gravity (NSG):

```{r wood-bio}
# Get NSG (wood density) from lookup - lookupcode converts species name to the chosen code
nsg_lookup <- lookupcode("Oak", code = "NSG")
nsg <- nsg_lookup$code

# Calculate wood biomass
wood_bio <- woodbiomass(
  treevol = stem_vol$stemvolume,
  nsg = nsg,
  sig_treevol = stem_vol$sigma,
  sig_nsg = 0.094  # Default NSG uncertainty
)

print(wood_bio)
```

## 4.2 Crown Biomass (Equations 6-7)

Crown biomass is estimated from species-specific crown factors:

```{r crown-bio}
# Calculate crown biomass
crown_bio <- crownbiomass(
  spcode = "OK",
  dbh = 45,
  re_d = 0.025,
  re = 0.025
)

print(crown_bio)
```

## 4.3 Root Biomass (Equations 8-9)

Below-ground biomass:

```{r root-bio}
# Calculate root biomass
root_bio <- rootbiomass(
  spcode = "OK",
  dbh = 45,
  re_dbh = 0.025,
  re = 0.025
)

print(root_bio)
```

## 4.4 Total Above-Ground Biomass

```{r agb}
# Sum wood and crown biomass
agb <- wood_bio$woodbiomass + crown_bio$biomass
agb_sigma <- sqrt(wood_bio$sigma^2 + crown_bio$sigma^2)

cat(sprintf("Above-ground biomass: %.3f ± %.3f t\n", agb, agb_sigma))
```

---

# 5: Carbon Conversion

## 5.1 Biomass to Carbon

**Matthews2** is recommended for WCC (Matthews 1993, UK Forestry Commission).
It is the default in `fc_agc()` and `fc_agc_error()`. For uncertainty propagation
through carbon conversion, use IPCC2 or Thomas.

```{r biomass2c}
# Matthews2 (recommended for WCC) - type-specific carbon fraction
carbon_matthews2 <- biomass2c(biomass = agb, method = "Matthews2", type = "broadleaf")
print(carbon_matthews2)

# Thomas or IPCC2 - for uncertainty propagation (includes sig_AGC)
carbon_thomas <- biomass2c(
  biomass = agb,
  method = "Thomas",
  type = "broadleaf",
  sig_biomass = agb_sigma
)
print(carbon_thomas)

# IPCC2 (alternative with uncertainty)
carbon_ipcc <- biomass2c(
  biomass = agb,
  method = "IPCC2",
  type = "broadleaf",
  sig_biomass = agb_sigma
)
print(carbon_ipcc)
```

---

# 6: Complete Pipeline - Step by Step

Here's the complete WCC pipeline for a single tree:

```{r complete-pipeline}
# === Step 1: Species lookup ===
species_name <- "Oak"
spcode_result <- lookupcode(species_name, type = "broadleaf")
spcode <- spcode_result$code
# Get NSG (wood density) when needed
nsg_lookup <- lookupcode(species_name, code = "NSG")
nsg <- nsg_lookup$code

cat("=== Step 1: Species Lookup ===\n")
cat(sprintf("Species: %s\n", species_name))
cat(sprintf("WCC Code: %s\n", spcode))
cat(sprintf("NSG: %s\n", nsg))
cat("\n")

# === Step 2: Tariff calculation ===
dbh <- 45  # cm
height <- 18  # m
tariff_result <- tariffs(
  spcode = spcode,
  height = height,
  dbh = dbh,
  type = "broadleaf",
  re_h = 0.05,
  re_dbh = 0.025
)

cat("=== Step 2: Tariff Calculation ===\n")
cat(sprintf("Tariff: %.2f ± %.2f\n", tariff_result$tariff, tariff_result$sigma))
cat("\n")

# === Step 3: Volume estimation ===
merch_vol <- merchtreevol(
  dbh = dbh,
  tariff = tariff_result$tariff,
  re_dbh = 0.025,
  sig_tariff = tariff_result$sigma
)

stem_vol <- treevol(
  mtreevol = merch_vol$volume,
  dbh = dbh,
  sig_mtreevol = merch_vol$sigma
)

cat("=== Step 3: Volume Estimation ===\n")
cat(sprintf("Merchantable volume: %.3f ± %.3f m³\n", 
            merch_vol$volume, merch_vol$sigma))
cat(sprintf("Stem volume: %.3f ± %.3f m³\n", 
            stem_vol$stemvolume, stem_vol$sigma))
cat("\n")

# === Step 4: Biomass calculation ===
# Get NSG (wood density) from lookup
nsg_lookup <- lookupcode(species_name, code = "NSG")
nsg <- nsg_lookup$code
wood_bio <- woodbiomass(
  treevol = stem_vol$stemvolume,
  nsg = nsg,
  sig_treevol = stem_vol$sigma,
  sig_nsg = 0.094
)

crown_bio <- crownbiomass(
  spcode = spcode,
  dbh = dbh,
  re_d = 0.025
)

agb <- wood_bio$woodbiomass + crown_bio$biomass
agb_sigma <- sqrt(wood_bio$sigma^2 + crown_bio$sigma^2)

cat("=== Step 4: Biomass Calculation ===\n")
cat(sprintf("Wood biomass: %.3f ± %.3f t\n", 
            wood_bio$woodbiomass, wood_bio$sigma))
cat(sprintf("Crown biomass: %.3f ± %.3f t\n", 
            crown_bio$biomass, crown_bio$sigma))
cat(sprintf("Total AGB: %.3f ± %.3f t\n", agb, agb_sigma))
cat("\n")

# === Step 5: Carbon conversion ===
# Using Thomas for uncertainty propagation in this demo; Matthews2 is WCC default
carbon <- biomass2c(
  biomass = agb,
  method = "Thomas",
  type = "broadleaf",
  sig_biomass = agb_sigma
)

cat("=== Step 5: Carbon Conversion ===\n")
cat(sprintf("Above-ground carbon: %.3f ± %.3f t C\n", 
            carbon$AGC, carbon$sig_AGC))
cat("\n")

# === Summary ===
cat("=== FINAL RESULT ===\n")
cat(sprintf("Tree: %s (DBH: %d cm, Height: %d m)\n", species_name, dbh, height))
cat(sprintf("Above-ground carbon: %.3f ± %.3f t C\n", 
            carbon$AGC, carbon$sig_AGC))
cat(sprintf("95%% CI: [%.3f, %.3f] t C\n",
            carbon$AGC - 1.96 * carbon$sig_AGC,
            carbon$AGC + 1.96 * carbon$sig_AGC))
```

---

# 7: Using the High-Level Functions

For most users, the high-level functions `fc_agc()` and `fc_agc_error()` are more convenient:

## 7.1 Simple Carbon Calculation

```{r fc-agc-simple}
# Simple calculation - returns carbon only
carbon_simple <- fc_agc(name = "Oak", dbh = 45, height = 18,
                        type = "broadleaf", output.all = FALSE)

cat(sprintf("Carbon: %.3f tonnes\n", carbon_simple))
```

## 7.2 Detailed Output

```{r fc-agc-detailed}
# Get all intermediate values
carbon_detailed <- fc_agc(name = "Oak", dbh = 45, height = 18,
                          type = "broadleaf", output.all = TRUE)

print(carbon_detailed)
```

## 7.3 With Error Propagation

```{r fc-agc-error}
# Include uncertainty estimates
carbon_error <- fc_agc_error(name = "Oak", dbh = 45, height = 18,
                             type = "broadleaf",
                             re_dbh = 0.025, re_h = 0.05, re = 0.025)

# View key results
cat("=== Carbon with Uncertainty ===\n")
cat(sprintf("Carbon: %.3f ± %.3f t C\n", 
            carbon_error$AGC_WCC_t, carbon_error$sig_AGC))
cat(sprintf("95%% CI: [%.3f, %.3f] t C\n",
            carbon_error$AGC_WCC_t - 1.96 * carbon_error$sig_AGC,
            carbon_error$AGC_WCC_t + 1.96 * carbon_error$sig_AGC))
```

## 7.4 Rich Output with Metadata

```{r fc-agc-rich}
# Get comprehensive metadata
carbon_rich <- fc_agc(name = "Oak", dbh = 45, height = 18,
                      type = "broadleaf", rich_output = TRUE)

print(carbon_rich)
```

---

# 8: Batch Processing Multiple Trees

## 8.1 Multiple Trees with fc_agc()

```{r batch-fc-agc}
# Create sample data
trees <- data.frame(
  name = c("Oak", "Beech", "Ash", "Birch"),
  dbh = c(45, 38, 52, 28),
  height = c(18, 22, 24, 15),
  type = rep("broadleaf", 4)
)

# Calculate carbon for all trees
results <- fc_agc(name = trees$name, dbh = trees$dbh,
                  height = trees$height, type = trees$type,
                  output.all = TRUE)

# Summary
cat("=== Batch Results ===\n")
cat(sprintf("Total carbon: %.2f tonnes\n", sum(results$AGC_WCC_t, na.rm = TRUE)))
cat(sprintf("Mean per tree: %.3f tonnes\n", mean(results$AGC_WCC_t, na.rm = TRUE)))
print(results[, c("name", "dbh", "height", "AGC_WCC_t")])
```

## 8.2 Batch Processing with Error Propagation

```{r batch-fc-agc-error}
# With uncertainty
results_error <- fc_agc_error(name = trees$name, dbh = trees$dbh,
                              height = trees$height, type = trees$type,
                              re_dbh = 0.025, re_h = 0.05, re = 0.025)

# Calculate total uncertainty (error propagation for sum)
total_carbon <- sum(results_error$AGC_WCC_t, na.rm = TRUE)
total_uncertainty <- sqrt(sum(results_error$sig_AGC^2, na.rm = TRUE))

cat("=== Batch Results with Uncertainty ===\n")
cat(sprintf("Total carbon: %.2f ± %.2f tonnes\n", total_carbon, total_uncertainty))
cat(sprintf("95%% CI: [%.2f, %.2f] tonnes\n",
            total_carbon - 1.96 * total_uncertainty,
            total_carbon + 1.96 * total_uncertainty))
```

## 8.3 Batch Processing with Rich Output

```{r batch-rich}
# Rich output for multiple trees
results_rich <- fc_agc(name = trees$name, dbh = trees$dbh,
                       height = trees$height, type = trees$type,
                       rich_output = TRUE)

print(results_rich)
```

---

# 9: Seedlings and Saplings

For trees with DBH < 7 cm or height < 10 m, use the seedling/sapling model.
**WCC definitions:** Seedlings = height < 1 m OR DBH < 7 cm; Saplings = height 1–10 m
OR DBH 7–10 cm; Trees = DBH >= 7 cm AND height >= 10 m. Height is in **metres**:

```{r seedlings}
# Seedling (height 0.5 m = 50 cm)
seedling_carbon <- sap_seedling2C(
  height = 0.5,     # metres (50 cm)
  type = "broadleaf",
  re_h = 0.05,
  re = 0.025
)

cat(sprintf("Seedling carbon: %.4f ± %.4f t C\n",
            seedling_carbon$carbon, seedling_carbon$sd))

# Sapling (height 5 m)
sapling_carbon <- sap_seedling2C(5, "conifer")
cat(sprintf("Sapling carbon: %.4f t C\n", sapling_carbon))
```

---

# 10: Custom Wood Density

You can specify your own wood density (NSG) instead of using WCC defaults:

```{r custom-nsg}
# Get wood density from BIOMASS package
if (requireNamespace("BIOMASS", quietly = TRUE)) {
  wd <- BIOMASS::getWoodDensity("Quercus", "robur", region = "Europe")
  
  # Use custom NSG
  carbon_custom <- fc_agc(name = "Oak", dbh = 45, height = 18,
                          type = "broadleaf", nsg = wd$meanWD)
  
  cat("=== Using Custom Wood Density ===\n")
  cat(sprintf("NSG from BIOMASS: %.3f\n", wd$meanWD))
  cat(sprintf("Carbon: %.3f tonnes\n", carbon_custom$AGC_WCC_t))
}
```

---

# 11: Visualization Examples

## 11.1 Carbon vs DBH

```{r plot-dbh, fig.height=5}
# Create data for plotting
dbh_range <- seq(10, 100, by = 5)
carbon_result <- fc_agc(name = rep("Oak", length(dbh_range)),
                        dbh = dbh_range,
                        height = rep(18, length(dbh_range)),
                        type = rep("broadleaf", length(dbh_range)),
                        output.all = FALSE)

# Extract carbon values from result (handles both data frame and vector return)
if (is.data.frame(carbon_result)) {
  carbon_range <- carbon_result$AGC_WCC_t
} else {
  carbon_range <- carbon_result
}

plot_data <- data.frame(
  dbh = dbh_range,
  carbon = carbon_range
)

ggplot(plot_data, aes(x = dbh, y = carbon)) +
  geom_line(linewidth = 1.2, color = "darkgreen") +
  geom_point(size = 2, color = "darkgreen") +
  labs(title = "Above-Ground Carbon vs DBH (Oak, Height = 18m)",
       x = "DBH (cm)", y = "Above-Ground Carbon (tonnes)") +
  theme_minimal()
```

## 11.2 Carbon with Uncertainty

```{r plot-uncertainty, fig.height=5}
# Calculate with uncertainty
dbh_range <- seq(20, 80, by = 5)
results_unc <- fc_agc_error(name = rep("Oak", length(dbh_range)),
                            dbh = dbh_range,
                            height = rep(18, length(dbh_range)),
                            type = rep("broadleaf", length(dbh_range)),
                            re_dbh = 0.025, re_h = 0.05)

plot_data <- data.frame(
  dbh = dbh_range,
  carbon = results_unc$AGC_WCC_t,
  ci_low = results_unc$AGC_WCC_t - 1.96 * results_unc$sig_AGC,
  ci_high = results_unc$AGC_WCC_t + 1.96 * results_unc$sig_AGC
)

ggplot(plot_data, aes(x = dbh, y = carbon)) +
  geom_ribbon(aes(ymin = ci_low, ymax = ci_high), 
              alpha = 0.2, fill = "darkgreen") +
  geom_line(linewidth = 1.2, color = "darkgreen") +
  geom_point(size = 2, color = "darkgreen") +
  labs(title = "Above-Ground Carbon vs DBH with 95% Confidence Intervals",
       x = "DBH (cm)", y = "Above-Ground Carbon (tonnes)") +
  theme_minimal()
```

---

# 12: Common Workflows

## 12.1 Method D: Stand-Level Carbon (fc_stand_carbon)

For stand-level assessment when you have DBH and height from sample trees,
use `fc_stand_carbon()` (WCC Method D). It calculates top height and mean
DBH internally:

```{r method-d-stand-carbon}
# Method D: Stand-level carbon from sample trees
dbh_stand <- c(45, 52, 38, 60, 42, 55, 48, 35, 58, 40)
height_stand <- c(18, 22, 15, 24, 17, 21, 19, 14, 23, 16)
area_ha <- 0.1

stand_carbon <- fc_stand_carbon(name = "Oak", dbh = dbh_stand,
                                height = height_stand, area_ha = area_ha,
                                type = "broadleaf",
                                re_dbh = 0.05, re_h = 0.05,
                                output.all = TRUE)

cat("=== Method D: Stand-Level Carbon ===\n")
print(stand_carbon)
```

## 12.2 Method E: DBH Class Tallies (wcc_process_tally, wcc_dbh_class_carbon)

When trees are tallied by DBH class rather than measured individually
(WCC Method E), use `wcc_process_tally()` then `wcc_dbh_class_carbon()`:

```{r method-e-dbh-class}
# Method E: DBH class tallies
dbh_classes <- c(5, 15, 25, 35, 45)   # cm midpoints
counts <- c(150, 200, 180, 120, 50)   # trees per class
heights <- c(3, 8, 15, 20, 22)        # representative heights (m)

tally_data <- wcc_process_tally(dbh_class = dbh_classes, count = counts,
                                height = heights, use_midpoints = TRUE)

# Carbon from tally
carbon_tally <- wcc_dbh_class_carbon(name = "Oak", tally_data = tally_data,
                                     type = "broadleaf", area_ha = 1.0,
                                     re_dbh = 0.05, re_h = 0.05)

cat("=== Method E: DBH Class Carbon ===\n")
print(carbon_tally)
```

## 12.3 Forest Stand Carbon Inventory (Methods B/C)

```{r stand-inventory}
# Example: Forest stand with multiple species
stand_data <- data.frame(
  species = c("Oak", "Beech", "Ash", "Birch", "Oak", "Beech"),
  dbh = c(45, 38, 52, 28, 55, 42),
  height = c(18, 22, 24, 15, 20, 21),
  type = c("broadleaf", "broadleaf", "broadleaf", "broadleaf", 
           "broadleaf", "broadleaf")
)

# Calculate carbon
stand_results <- fc_agc_error(name = stand_data$species, dbh = stand_data$dbh,
                              height = stand_data$height,
                              type = stand_data$type,
                              re_dbh = 0.025, re_h = 0.05)

# Summary by species
stand_summary <- aggregate(
  list(AGC = stand_results$AGC_WCC_t,
       Uncertainty = stand_results$sig_AGC),
  by = list(Species = stand_results$name),
  FUN = function(x) c(Total = sum(x, na.rm = TRUE), 
                      Mean = mean(x, na.rm = TRUE))
)

cat("=== Stand Carbon Inventory ===\n")
print(stand_summary)

# Total stand carbon
total_stand_carbon <- sum(stand_results$AGC_WCC_t, na.rm = TRUE)
total_stand_uncertainty <- sqrt(sum(stand_results$sig_AGC^2, na.rm = TRUE))

cat(sprintf("\nTotal Stand Carbon: %.2f ± %.2f tonnes C\n",
            total_stand_carbon, total_stand_uncertainty))
```

## 12.4 Carbon per Unit Area

```{r per-area}
# Calculate carbon density (tonnes C per hectare)
stand_area_ha <- 0.5  # 0.5 hectares

carbon_density <- total_stand_carbon / stand_area_ha
carbon_density_uncertainty <- total_stand_uncertainty / stand_area_ha

cat("=== Carbon Density ===\n")
cat(sprintf("Stand area: %.2f ha\n", stand_area_ha))
cat(sprintf("Carbon density: %.2f ± %.2f t C/ha\n",
            carbon_density, carbon_density_uncertainty))
```

## 12.5 WCC Averaging and Scaling Helpers

The WCC protocol often requires averaging trees (e.g., mean tariff number) and scaling
to per-hectare values. Helper functions support these operations:

```{r wcc-mean-tariff}
# Mean tariff from sample trees (WCC protocol requirement)
spcode <- lookupcode("Oak", type = "broadleaf")$code
sample_heights <- c(18, 20, 19, 22, 17)  # m
sample_dbh <- c(45, 48, 42, 52, 38)      # cm

mean_tariff <- wcc_mean_tariff(
  spcode = spcode,
  height = sample_heights,
  dbh = sample_dbh,
  type = "broadleaf",
  re_h = 0.05,
  re_dbh = 0.025
)

cat("=== Mean Tariff from Sample Trees ===\n")
if (is.list(mean_tariff)) {
  cat(sprintf("Mean tariff: %.2f ± %.2f\n", mean_tariff$mean_tariff, mean_tariff$sigma))
} else {
  cat(sprintf("Mean tariff: %.2f\n", mean_tariff))
}
```

```{r wcc-per-hectare}
# Scale per-tree carbon to per-hectare
carbon_per_tree <- c(0.5, 0.6, 0.45, 0.55, 0.5)  # t C per tree
trees_per_ha <- 200
plot_area_ha <- 0.5

per_ha_result <- wcc_per_hectare(
  carbon_per_tree = carbon_per_tree,
  trees_per_ha = trees_per_ha,
  area_ha = plot_area_ha,
  sig_carbon_per_tree = 0.05  # optional: propagates uncertainty
)

cat("=== Per-Hectare Scaling ===\n")
cat(sprintf("Carbon per ha: %.2f ± %.2f t C/ha\n", 
            per_ha_result$carbon_per_ha, per_ha_result$sig_carbon_per_ha))
cat(sprintf("Total carbon (%.1f ha): %.2f ± %.2f t C\n", 
            plot_area_ha, per_ha_result$total_carbon, per_ha_result$sig_total_carbon))
```

```{r wcc-stratify}
# Aggregate carbon across stratified stands (with optional uncertainty)
strata <- data.frame(
  area_ha = c(0.5, 0.3, 0.2),
  carbon_per_ha_t = c(120, 95, 80),
  carbon_per_ha_sig_t = c(15, 12, 10)  # optional: per-stratum uncertainty
)

strata_result <- wcc_stratify(strata, carbon_col = "carbon_per_ha_t")

cat("=== Stratified Carbon Aggregation ===\n")
cat(sprintf("Total area: %.2f ha\n", strata_result$total_area_ha))
cat(sprintf("Total carbon: %.1f ± %.1f t C\n", 
            strata_result$total_carbon_t, strata_result$total_sig_carbon_t))
cat(sprintf("Weighted mean: %.1f ± %.1f t C/ha\n", 
            strata_result$weighted_mean_carbon_per_ha_t, strata_result$weighted_mean_sig))
```

---

# 13: Troubleshooting

## 13.1 Species Not Found

If a species isn't in the WCC database:

```{r troubleshooting-species}
# Unknown species - will use type-based defaults
result_unknown <- fc_agc(name = "Unknown Species", dbh = 45, height = 18,
                         type = "broadleaf")  # Important: specify type

cat("Using type-based defaults for unknown species\n")
print(result_unknown)
```

## 13.2 Invalid DBH/Height Ranges

```{r troubleshooting-range}
# Very small tree - may trigger warnings
result_small <- fc_agc(name = "Oak", dbh = 5, height = 3,
                       type = "broadleaf")

# Check for warnings/flags
cat("Small tree calculation (may have validity warnings)\n")
```

## 13.3 Missing Height

For WCC, height is required. If you don't have height:

```{r troubleshooting-height}
# Option 1: Use Bunce method instead (no height needed)
bunce_result <- Bunce(name = "Oak", dbh = 45)

# Option 2: Estimate height from DBH (not recommended for WCC)
# WCC requires measured height for accurate carbon accounting
```

---

# 14: Best Practices

1. **Always specify type** for ambiguous species names
2. **Use measured height** - WCC requires actual height measurements
3. **Include uncertainty** - Use `fc_agc_error()` for proper error propagation
4. **Use Matthews2 for carbon** - Default in WCC functions; recommended by WCC protocol (Matthews 1993)
5. **Use appropriate methods** - WCC is for UK temperate woodland
6. **Validate inputs** - Ensure DBH and height are in correct units (cm and m)

---

# References

1. **Woodland Carbon Code**: Jenkins, T.A.R., Mackie, E.D., Matthews, R.W., Miller, G., Randle, T.J., & White, M.E. (2018). FC Woodland Carbon Code: Carbon Assessment Protocol (v2.0). Forestry Commission, Edinburgh. https://www.woodlandcarboncode.org.uk/

---

# Session Info

```{r session-info}
sessionInfo()
```
